---
title: "IQ Analysis"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



###################################
 
## Stat Consulting: Hw 2
Author: Erika Esquinca
Class:  Bios 6621
        Dr. Ryan Peterson
Date:   9.15.2020


The purpose of this assignment is to be able to create a new project in R using version
control and a git repository. 
This makes me work reproducible as any changes, previous code, and directories are all
linked to github to store on the cloud.

This report is on IQ levels based on lead exposure and must contain the following.


A. A graph showing the IQ levels by location status

B. At least one nicely formatted table using the 'kable' function

C. A couple sentences of text (not on the graph, but in the body of the Rmarkdown document) describing the graph and the table

D. Inline calculations giving the values of the means, that will be updated if the data are changed,

E. At least one R code chunk.


First I will begin by exporting the data and taking a look at some values

```{r}
iq <- read.csv("lead-iq-01.csv")
```


Lets begin the analyses by looking at the first 6 rows of our data and the structure

```{r}
str(iq)
head(iq)
```
Looks like Smelter is a factor with two level "Far" = 1 & "Near" = 0 and IQ is an integer value.


A. A graph showing the IQ levels by location status

```{r}
library(ggplot2)
qplot(x= iq$Smelter, y= iq$IQ, 
      xlab= "Smelter",
      ylab= "IQ",
      main= "Ranges of IQ when far or near smelter")
```

We can also interpret this data using boxplots, however notice that our outputs are 
hard to read because of the outlier changing the scale

```{r}
plot(x= iq$Smelter, y= iq$IQ, 
      xlab= "Smelter",
      ylab= "IQ",
      main= "Ranges of IQ when far or near smelter")

```


B. At least one nicely formatted table using the 'kable' function

```{r}
library(knitr)
kable(summary(iq))
kable(summary(iq[iq$Smelter == "Far",]))
kable(summary(iq[iq$Smelter == "Near",]))

```


C. A couple sentences of text (not on the graph, but in the body of the Rmarkdown document) describing the graph and the table

Notice now that the graphs ahow the IQ results obtained from individuals that are either considered near smelter which is a subject being within 1 mile of lead emitting ore smelter or more than 1 mile considered far. The most notable observation is that in the far category we have a subject with an IQ of 999.
This then means we have a larger range of possible IQ's for the far category. 

D. Inline calculations giving the values of the means, that will be updated if the data are changed,

We can also now input values we can see using inline R code that get updated when the document
gets knitted. This way if data changes you don't have to worry about integrating correct values into
the text. 

An example of this is the mean of the IQ data we have is `r mean(iq$IQ)`


 








